---
title: "Mini Literate Analysis"
format: html
---

## Introduction of my analysis

### About the data

This is chemistry data. The data was collected from the ocean using water bootles. This data was downloaded from the following link of Arctic Data Center [Arctic Data Center](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X).

### Setup of the analysis

```{r}

library(readr)

```

### Let's read the data!

```{r}

bg_chem <- read_csv("data/BGchem2008data.csv")

```

Exploring data

```{r}

colnames(bg_chem)
summary(bg_chem)

#Prints all the dates in a specific column:

unique(bg_chem$Date)

```

## Analysis

### Let's calculate summary statistics

```{r}

nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)

```

### Let's calculate the mean Readfield ratio

```{r}

ratio <- (nitrate + nitrite + amm)/phos

```

### Let's plot the ratio

```{r}

plot(bg_chem$P,bg_chem$NO3 + bg_chem$NO2 + bg_chem$NH4)

```

## Conclusion

The Redfield ratio is `r round(ratio)`

